---
title: "Lab10"
author: "VY"
date: "`r Sys.Date()`"
output: 
  github_document: 
  html_document: default
  html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com")) #install packages when knitting

# check if packages installed yet
if(!require(RSQLite)) install.packages("RSQLite")
if(!require(DBI)) install.packages("DBI")

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)

dbListTables(con)

dbGetQuery(con, "PRAGMA table_info(actor)")
```

```{sql connection=con}
PRAGMA table_info(actor)
```


## Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
```

Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{sql connection=con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER by last_name, first_name
```

Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{sql connection=con}
SELECT DISTINCT rental_id
FROM rental
WHERE date(rental_date) = '2005-07-05'
```

Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{sql connection=con}
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
```
Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{sql connection=con}
SELECT *
FROM payment
WHERE amount > 5
```


```{r clean up}
# clean up
dbDisconnect(con)
```

