---
title: "Lab10"
author: "VY"
date: "`r Sys.Date()`"
output: 
  github_document: 
  html_document: default
  html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com")) #install packages when knitting

# check if packages installed yet
if(!require(RSQLite)) install.packages("RSQLite")
if(!require(DBI)) install.packages("DBI")

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)

dbListTables(con)

dbGetQuery(con, "PRAGMA table_info(actor)")
```

```{sql connection=con}
PRAGMA table_info(actor)
```


## Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
```

Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{sql connection=con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER by last_name, first_name
```

Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{sql connection=con}
SELECT DISTINCT customer_id, rental_date
FROM rental
WHERE date(rental_date) = '2005-07-05'
```

Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{sql connection=con}
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
```
Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{sql connection=con}
SELECT *
FROM payment
WHERE amount > 5
```
Exercise 4.3
Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8
```{sql connection=con}
SELECT *
FROM payment
WHERE amount > 5 AND amount < 8
```
Exercise 5
Retrive all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.

```{sql connection=con}
SELECT payment_id, amount, last_name, first_name, a.customer_id
FROM payment AS a
  INNER JOIN customer AS b
  ON a.customer_id = b.customer_id
WHERE b.last_name in ('DAVIS')
```

Exercise 6.1
Use COUNT(*) to count the number of rows in rental
```{sql connection=con}
SELECT 
  COUNT(*) AS N
FROM rental
```

Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{sql connection=con}
SELECT customer_id,
  COUNT(*) AS count
FROM rental
GROUP BY customer_id
```

Exercise 6.3
Repeat the previous query and sort by the count in descending order
```{sql connection=con}
SELECT customer_id,
  COUNT(*) AS count
FROM rental
GROUP BY customer_id
ORDER BY count DESC
```

Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{sql connection=con}
SELECT customer_id,
  COUNT(*) AS count
FROM rental
GROUP BY customer_id
HAVING count >= 40
ORDER BY count DESC
```


Exercise 7
The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM
```{sql connection=con}
SELECT MAX(amount) AS maxamount, MIN(amount) AS minamount, AVG(amount) AS avgamount, SUM(amount) AS sumamount
FROM payment
```

Exercise 7.1
Modify the above query to do those calculations for each customer_id
```{sql connection=con}
SELECT customer_id, MAX(amount) AS maxamount, MIN(amount) AS minamount, AVG(amount) AS avgamount, SUM(amount) AS sumamount
FROM payment
GROUP BY customer_id
```
Exercise 7.2
Modify the above query to only keep the customer_ids that have more then 5 payments
```{sql connection=con}
SELECT customer_id, COUNT(*) AS N, MAX(amount) AS maxamount, MIN(amount) AS minamount, AVG(amount) AS avgamount, SUM(amount) AS sumamount
FROM payment
GROUP BY customer_id
HAVING N >5
```

```{r clean up}
# clean up
dbDisconnect(con)
```

