---
title: "Lab04"
author: "VY"
date: "`r Sys.Date()`"
output: html_document
---
Download file if needed and remove any temp less than -17 and look for key missing data. No 9999 in dew point or wind speed.
```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../Lab04/met_all.gz")

met <- met[temp>-17][elev == 9999.0, elev := NA]

str(met)

summary(met$elev)
summary(met$temp)
summary(met$dew.point)
summary(met$wind.sp)

library(tidyverse)
library(data.table)
```


New date column, keep only first week data
```{r}
met <- met[ , ymd := as.Date(paste(year, month, day, sep= "-"))]
table(met$ymd)
met[, table(week(ymd))]
met <- met[week(ymd)==31]
nrow(met)
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev. (no 9999s)
```{r}
met[, .(
  temp = max(temp),
  rh = max(rh, na.rm=T),
  wind.sp = max(wind.sp, na.rm=T),
  vis.dist = max(vis.dist, na.rm=T),
  dew.point = max(dew.point, na.rm=T),
  lat = max(lat),
  lon = max(lon),
  elev = max(elev, na.rm=T)
)]
```

```{r}
met_avg <- met[, .(
  temp = mean(temp),
  rh = mean(rh, na.rm=T),
  wind.sp = mean(wind.sp, na.rm=T),
  vis.dist = mean(vis.dist, na.rm=T),
  dew.point = mean(dew.point, na.rm=T),
  lat = mean(lat),
  lon = mean(lon),
  elev = mean(elev, na.rm=T)
), by = "USAFID"]
head(met_avg)
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}

```

Create a categorical variable for elevation as in the lecture slides
```{r}

```

