---
title: "Lab05"
author: "VY"
date: '`r Sys.Date()`'
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read-data install packages, cache=T}

webshot::install_phantomjs()
library(leaflet)
library(webshot)
library(data.table)
library(dtplyr)
library(dplyr)

if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../Lab04/met_all.gz")
met <- met[temp>-17][elev == 9999.0, elev := NA]
```

```{r read in new station data, cache=T}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r merge data}
met <- merge(
  x = met,
  y = stations,
  by.x = "USAFID",
  by.y = "USAF",
  all.x = TRUE,
  all.y = FALSE
)
```

#What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?
```{r mean median country}
station_averages <- 
  met[ , .(
    temp=mean(temp, na.rm=T),
    wind.sp=mean(wind.sp, na.rm=T),
    atm.press=mean(atm.press, na.rm=T)
  ), by = "USAFID"]
stat_medians <- station_averages[ , .(
    temp50 =median(temp, na.rm=T),
    wind.sp50 =median(wind.sp, na.rm=T),
    atm.press50 =median(atm.press, na.rm=T)
  )]
station_averages[ , temp_dist50 := abs(temp - stat_medians$temp50)][order(temp_dist50)]
station_averages[which.min(temp_dist50)]
```

#Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.

```{r most representative state}
state_averages <- 
  met[ , .(
    temp=mean(temp, na.rm=T),
    wind.sp=mean(wind.sp, na.rm=T),
    atm.press=mean(atm.press, na.rm=T)
  ), by = .(USAFID,STATE)]
state_medians <- state_averages[ , .(
    temp50 =median(temp, na.rm=T),
    wind.sp50 =median(wind.sp, na.rm=T),
    atm.press50 =median(atm.press, na.rm=T)
  ), by = STATE]

statemerge <-merge( 
  x = state_averages,
  y=state_medians,
  by.x = "STATE",
  by.y = "STATE",
  all.x=TRUE,
  all.y=FALSE)

station_averages[ , tempstatedist_50 := temp - statemerge$temp50]
station_averages[ , windstatedist_50 := wind.sp - statemerge$wind.sp50]
station_averages[ , eucdist := tempstatedist_50^2 + windstatedist_50^2 ]

station_averages[ which.min(eucdist)]
                    
```
Station 726625
