---
title: "Lab05"
author: "VY"
date: '`r Sys.Date()`'
output: github_document
always_allow_html: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up packages, cache = TRUE}
library(dplyr)
library(data.table)
library(ggplot2)
library(tidytext)
library(tidyverse)
library(dtplyr)
library(knitr)
library(forcats)
library(utils)
```
## Read in data
```{r read data, cache=TRUE}
if (!file.exists("mtsamples.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
    )
mtsamples <- read.csv("mtsamples.csv")
mts <- as_tibble(mtsamples)
```
## What specialties do we have? What are the frequencies of each?
```{r}
specialties <- count(mts, medical_specialty)

specialties %>%
    arrange(desc(n)) %>% 
      knitr::kable()
```
There are `r nrow(specialties)` medical specialties. The frequency is not evenly distributed. Surgery is most common with 1103 entries, Hospice - Palliative Care is the least frequent with 6 entries. Some are related - ie Consult - History and Physical and Discharge Summary are not specialty-specific. 
```{r graph specialties}
specialties %>% 
  top_n(10) %>%
  ggplot(aes(x=n, y= fct_reorder(medical_specialty, n))) +
  geom_col()
```
## Tokenize words in transcription and look at top 20
```{r tokenize, cache = TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort=TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()
```
There are a lot of stop words here. The top word is "the" while we do see "patient". We should get rid of the stop words.

## Remove stop words
```{r remove stopwords}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort=TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x=word)) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()
  
```
These look more likel medical terms. Numbers were removed from the list of words.
